### 2018-08-13 paper/vis.tex

% $Id: template.tex 11 2007-04-03 22:25:53Z jpeltier $

%\documentclass{vgtc}                          % final (conference style)
\documentclass[review]{vgtc}                 % review
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint]{vgtc}               % preprint
%\documentclass[electronic]{vgtc}             % electronic version

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Figures should be in CMYK or Grey scale format, otherwise, colour 
%% shifting may occur during the printing process.

%% These few lines make a distinction between latex and pdflatex calls and they
%% bring in essential packages for graphics and font handling.
%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
%% to provide the the path and extension of a graphics file:
%% \includegraphics{diamondrule} is completely sufficient.
%%
\ifpdf%                                % if we use pdflatex
  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
  \pdfcompresslevel=9                  % PDF Compression
  \pdfoptionpdfminorversion=7          % create PDF 1.7
  \ExecuteOptions{pdftex}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
\else%                                 % else we use pure latex
  \ExecuteOptions{dvips}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
\fi%

%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
\graphicspath{{figures/}{pictures/}{images/}{./}} % where to search for the images

\newcommand{\squishlist}{
 \begin{list}{$\bullet$}
  { \setlength{\itemsep}{0pt}
     \setlength{\parsep}{3pt}
     \setlength{\topsep}{3pt}
     \setlength{\partopsep}{0pt}
     \setlength{\leftmargin}{1.5em}
     \setlength{\labelwidth}{1em}
     \setlength{\labelsep}{0.5em} } }

\newcommand{\squishlisttwo}{
 \begin{list}{$\bullet$}
  { \setlength{\itemsep}{0pt}
    \setlength{\parsep}{0pt}
    \setlength{	opsep}{0pt}
    \setlength{\partopsep}{0pt}
    \setlength{\leftmargin}{2em}
    \setlength{\labelwidth}{1.5em}
    \setlength{\labelsep}{0.5em} } }

\newcommand{\squishend}{
  \end{list}  }

\usepackage{algpseudocode} 
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
\usepackage{textcomp}                  % use better special symbols
\usepackage{mathptmx}                  % use matching math font
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{cite}                      % needed to automatically sort the references
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.
\usepackage{color}
\newcommand{\G}{\ensuremath{G}}
\newcommand{\V}{\ensuremath{V}}
\newcommand{\vertice}{\ensuremath{v}}
\newcommand{\weight}{\ensuremath{w}}
\newcommand{\E}{\ensuremath{E}}
\newcommand{\Hist}{\ensuremath{H}}
\newcommand{\D}{\ensuremath{D}}
\newcommand{\relevance}{\ensuremath{R}}
\newcommand{\cluster}{\ensuremath{\mathcal{C}}}
\newcommand{\watershed}{\ensuremath{\psi}}

\newcommand{\lookDan}[1]{\marginpar{---}{\color{blue}{#1}}}

\DeclareMathOperator*{\median}{\ensuremath{\text{median}}}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{1101}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Visualizing demographic evolution using geographically inconsistent census data}

%% This is how authors are specified in the conference style

%% Author and Affiliation (single author).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}}
%%\affiliation{\scriptsize Allied Widgets Research}

%% Author and Affiliation (multiple authors with single affiliations).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
%%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
%%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
%%\affiliation{\scriptsize Martha Stewart Enterprises \\ Microsoft Research}

\author{Fabio Dias\thanks{e-mail: fabio.dias@utoronto.ca} %
\and Daniel Silver\thanks{e-mail: dsilver@utsc.utoronto.ca}}
\affiliation{\scriptsize University of Toronto}

%% A teaser figure can be included as follows, but is not recommended since
%% the space is now taken up by a full width abstract.
\teaser{
  \centering
 \includegraphics[width=0.85\linewidth]{newTeaser.pdf}
 \caption{Higher increase in the education level (purple cluster) in Chicago
  between 1970 and 2010. While the whole city follows this trend, the change was
  far more pronounced in these regions. The relevant clusters defined by Black
  (green) and White (orange) majority of population are also visible. 
  \label{fig:teaser}}}

%% Abstract section.
\abstract{Census measurements provide reliable demographic data going back
centuries. However, their analysis is often hampered by the lack of geographical
consistency across time. We propose a visual analytics system that enables the
exploration of geographically inconsistent data. Our method also includes
incremental developments in the representation, clustering, and visual
exploration of census data, allowing an easier  understanding of the demographic
groups present in a city and their evolution over time. We present the feedback
of experts in urban sciences and sociology, along with illustrative scenarios in
the USA and Canada.}

  % end of abstract

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/about/class> for details.
%% We recommend the 2012 system <http://www.acm.org/about/class/class/2012>
%% For the 2012 system use the ``\CCScatTwelve'' which command takes four arguments.
%% The 1998 system <http://www.acm.org/about/class/class/2012> is still possible
%% For the 1998 system use the ``\CCScat'' which command takes four arguments.
%% In both cases the last two arguments (1998) or last three (2012) can be empty.

\CCScatlist{ %TODO
  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visualization application domains}{Visual analytics};
  \CCScatTwelve{Mathematics of computing}{Probability and statistics}{Statistical paradigms}{Exploratory data analysis
  }
}

%\CCScatlist{
  %\CCScat{H.5.2}{User Interfaces}{User Interfaces}{Graphical user interfaces (GUI)}{};
  %\CCScat{H.5.m}{Information Interfaces and Presentation}{Miscellaneous}{}{}
%}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{lineno}
\linenumbers
\begin{document}
%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}
\maketitle
\input{intro}
%% \section{Introduction} %for journal use above \firstsection{..} instead

\input{related}

\input{design}

\input{method}

\input{study}

\input{discussion}

\section{Conclusions}
Our objective was to allow for the exploration of census data without
geographical normalization, an approach that is usually not considered:
\emph{"(...) tract-by-tract comparison is not possible unless data from 2000 is
interpolated to 2010 boundaries (...)"}~\cite{Dmowska2017}. Our method was able
to corroborate previous findings from the specialized literature, with an
increased level of detail due to our data representation and visualization
choices. The feedback from experts was remarkably positive, most of them were
able to extract insight from the prototype and demonstrated interest in using it
on their research efforts.

While we focused on census data in this work, the proposed method can be used
with most similarly area-based datasets. Further, while there is interest in the
use of census data, it is mostly in combination with other, larger and more
dynamic, datasets. We postulate that, while census data is interesting, the
amount of effort required to obtain and prepare it often discourages its use as
complementary information, reinforcing our hypothesis that there is need for
methods that are able to deal with geographically inconsistent data.

%% if specified like this the section will be ommitted in review mode
\acknowledgments{This research was supported by a University of Toronto
Connaught Global Challenge grant and is part of the Urban Genome Project.}
\bibliographystyle{abbrv-doi-hyperref-narrow}

\bibliography{rev}
\end{document}