### 2018-08-13 paper/study.tex
one past until LA

\section{Illustrative scenarios}
\label{sec:study}

We used decennial census data from the United States~\cite{nhgis} and
Canada\footnote{\url{http://datacentre.chass.utoronto.ca/census/}}, tabulated by
CTs,  from 1970 to 2010. The prototype allows access to 40 regions, 28 in the US
and 12 in Canada. Due to the high number of CTs, New York City was split into
its boroughs.

We used five aspects for the USA: Education level, Family income, Marital
status, Occupation, and Race; and seven for Canada: Age, Education level, Home
language, Household Income, Marital status, Occupation, Place of birth, and
Religion.  While our method does not require geographic harmonization, it
requires matching variables over time. The supplementary material contains the
details of which census columns were used for each aspect. Income is slightly
inaccurate, even though we did correct for official inflation. We grouped the
original ranges into three larger ranges, but they do not match precisely.

\subsection{Chicago}
We selected a region loosely following the administrative borders. The
demographic composition is well explored in the literature, with reports of
racial divide and gentrification~\cite{Delmelle2016,Delmelle2017,Hwang2014}.
While the definition of gentrification is still unclear and out of the scope of
this paper, we associate gentrification with higher education and income levels.

The initial state of the prototype is illustrated in Figure~\ref{fig:ui}, and
its findings are explained in Section~\ref{sec:ui}, where the racial divide is
clear. Starting from this initial state, the specific workflow used to identify
the existence and details of the gentrification process are illustrated in
Figure~\ref{fig:chiWorkflow}. For the users, the first step is to identify the
compositions of each cluster from the boxplots, so orange is associated with
majority of White population, green with majority Black, and purple with higher
proportion of four years of college or more (high education level). The expanded
version of the boxplots for the purple cluster shows a higher income level and
majority of occupations in administrative jobs, therefore the purple cluster
identifies gentrified regions.

The trajectories plot illustrates the process of gentrification, progressively
absorbing regions from the orange cluster (White). This corroborates results
from the literature reporting that Black neighborhoods are less likely to
gentrify~\cite{Hwang2014}. Moreover, this process is unlikely to be reversed, as
indicated by the limited number of trajectories leaving the purple stream. Next,
we select the region that is gentrified in 2010, by clicking on the
corresponding rectangle in the trajectories plot, updating the information on
the maps and the details portion of the interface.

The corresponding CTs are highlighted in the maps, where the spatial pattern is
clear, corresponding exactly to previous findings in the
literature~\cite{Hwang2014}. Further, we can also identify the regions that
gentrified earlier, with an stronger purple hue, compared to newer regions,
where the orange and green colors are still present. This also indicates from
which clusters they belonged before gentrifying. In the details portion of the
interface, the order of the aspects was updated to reflect the order of
relevance considering only the selected region. The most relevant aspect is the
Education, specifically "Four or more years of college", illustrated in the
rightmost portion of Figure~\ref{fig:chiWorkflow}, which is increasing for the
whole city (grey band), but faster and to a higher level in this region (black
band). Indeed, while the IQR for the city goes from 11\% to 45\%, the IQR for
this region goes from 55\% to 77\%. However, there are portions of this region
with significantly lower or higher proportions, as indicated by the dotted black
lines representing the minimum and maximum for the selected region.

\begin{figure*}
    \centering
 \includegraphics[width=0.85\linewidth]{newTeaser.pdf}
 \caption{Worflow to discover gentrification in Chicago: the purple cluster
 corresponds to high education / income. Its population is increasing over time,
 absorbing from the majority White cluster (orange). By selecting the purple
 cluster in 2010, the region is highlighted in the maps. The proportion of
 people with 4+ years of college is increasing in the whole city (grey IQRs),
 but significantly more in this region (black).\label{fig:chiWorkflow}}
\end{figure*}

\subsection{Toronto}

We considered a region that is approximately the administrative border of the
current city of Toronto and all seven available aspects with equal weights. The
results are summarized in Figure~\ref{fig:to}, considering eight clusters.

\begin{figure}
    \centering 
    \includegraphics[width=0.95\linewidth]{to.pdf}
    \caption{Clustering results for Toronto, with eight clusters, including
    clusters representing Jewish population, high and low income, low education,
    and Asian immigration.\label{fig:to}}
\end{figure}

The population with low percentage of University degrees is represented in
orange, mostly anglophone population in green, Asian immigrants in yellow, high
percentage of income in the highest bracket in purple, high percentage of Jewish
people in light green and brown, high percentage of Eastern Non-Christian
religion in pink, and high concentration of single people in dark gray. From the
trajectories plot, we can see that Toronto is more dynamic than Chicago, with
one cluster constantly shrinking. In the 1970s, the city was divided into four
clusters: low number of university degrees, Jewish population, majority
anglophones, and high income. Interestingly, the more recent clusters that
absorbed regions from the orange cluster have similar education profiles and are
differentiated by other aspects. In this sense, the city is growing diverse,
changing from a common low education profile to a higher level of education with
more diversity in religion (pink) and immigration (yellow).

Indeed, the influx of Asian population is visible starting in the 1980s and
building thereafter, leading to the yellow and pink clusters. While both include
a high percentage of people born in Asia, the pink is more defined by religion,
with low percentage of university degrees, and contains the lowest percentage of
people in the highest income bracket for these clusters; the yellow is less
defined by religion, and has higher education and income, geographically
corresponding to the Markham region, know for its Chinese population. A similar
division also happens for the two Jewish clusters, where the light green cluster
has lower education and income levels than the brown cluster. The purple cluster
of high income is somewhat stable. This cluster includes the Bridle Path
neighborhood, known for its wealthy population, until 2011. In 2011 it was
classified into the yellow cluster of Asian immigration, since about 35\% of the
population for this CT were then born in Asia. The income distribution did not
change, with 85\% of the population with an income of 90k CAD or more.

The most significant indicator of Toronto's dynamism is the presence of grey
regions on the simplified color map; representing regions associated to three or
more clusters over this five census period. Using the 'Add' mode for the
trajectory selection, we select their trajectories, and a subset of the details
is illustrated in Figure~\ref{fig:toVol}. These regions account for about 5\% of
Toronto's population. The whole region was classified into the orange cluster in
1971 (low level of university degrees). By 1991, most of the region was
classified into the green cluster, representing anglophone population, mostly
Canadian born, with a higher level of education. As the corresponding plot
indicates, this trend in increasing education is city-wide, but this region has
people with better education than most.

\begin{figure}
    \centering 
    \includegraphics[width=0.95\linewidth]{toVol.pdf}
    \caption{Details for some regions of Toronto that were classified into 3 or
         more clusters over time.\label{fig:toVol}}
\end{figure}

In 2001, the purple cluster of high income annexes neighboring parts of the
volatile region, and the Asian born population increases sharply, as illustrated
by the appearance of the yellow cluster.  This cluster indicates well educated,
higher income, and about 30\%-50\% Asian born population. By 2011, the yellow
cluster increased considerably, annexing parts of the high income purple
cluster, including the neighboring Bridle Path area.

\subsection{Los Angeles}%TODO HERE
We selected a region around the metropolitan area of Los Angeles (LA), following
urban density, not administrative boundaries. The summary of the results using
all aspects with equal weights and eight clusters is illustrated in
Figure~\ref{fig:la}. The full statistical description of the clusters is
illustrated in Figure~\ref{fig:laAll}, where the most relevant aspect of each
cluster is highlighted. From the trajectories plot, we can see that there is a
large but shrinking cluster, depicted in green, one increasing cluster in
purple, an almost constant orange cluster, a smaller but increasing pink
cluster, and three other small clusters. The corresponding map illustrates where
these clusters are located, and that they are somewhat geographically stable,
with some movement between the green, orange, and purple clusters.

\begin{figure}
    \centering 
    \includegraphics[width=0.95\linewidth]{LA.pdf}
    \caption{Result for Los Angeles with 8 clusters, including three small and
        ephemeral clusters. Cluster characterization is displayed
        in Figure~\ref{fig:laAll}.\label{fig:la}}
\end{figure}

From Figure~\ref{fig:laAll}, we can see that the green cluster is characterized by
a high percentage of White population, low percentage of population in the
lowest income bracket, mostly administrative occupations, and about 30\% of the
population with four or more years of college. The orange cluster is
characterized by a high percentage of Black population, with few people in the
highest range of income and education. The purple cluster corresponds to a high
concentration of "Other" in race, which includes Hispanic for this dataset, high
concentration of Laborers, and low education and income. The pink cluster
contains a high percentage of Asian population and about 30\% of the population
with four or more years of college. The light green cluster contains very few
people in the lower income bracket, mostly White population, with the highest
percentage of population with four or more years of college, working
administrative jobs, and a high concentration of singles. The yellow cluster
represents Black population, with higher level of education and income, mostly
working administrative jobs. The brown cluster represent a majority of single
population, working administrative jobs with mostly low income. The dark gray
cluster is characterized by all its population in the lowest income bracket, low
education level, with a majority of White population.

\begin{figure*}
    \centering 
    \includegraphics[width=0.95\linewidth]{laAll.pdf}
    \caption{Full characterization of the eight clusters found for LA. The red
     rectangles indicate the most relevant aspects for each
     cluster.\label{fig:laAll}}
\end{figure*}

Further, the larger clusters, green, orange, and purple, present a significant
intra-cluster variance in most variables, as indicated by the maximum and
minimum whiskers of the boxplots. For instance, while fifty percent of the CTs
in the green cluster have between 20\% and 40\% of people in the lowest income
bracket, that cluster also includes CTs where none and all the population
belongs to that bracket. This might indicate that this cluster could be further
divided, and currently represents different groups of people that are not
different enough to be separated into clusters in this level of the hierarchy.

While the larger clusters are expected, the light green, brown, and dark gray
clusters are more surprising. These clusters are small and ephemeral, including
only a few CTs, and are peculiarly different enough to be separated into their
own clusters, at this level of the hierarchy. While these small clusters might
be interesting enough to warrant further study, they might suppress larger, but
more subtle, clusters from appearing in this interface,  due to the limited
number of clusters.

The orange area in the map in Figure~\ref{fig:la} presents movement, indicated by
the presence of green and purple tones mixed with the orange, which may warrant
further exploration. By clicking on the orange bar above the boxplot, we select
all trajectories that contain the orange cluster. The corresponding details are
illustrated in Figure~\ref{fig:labh}. This shows a location change, where the
orange cluster is progressively replaced by the purple cluster on its east side,
and in turn expanding to the west. Interestingly, the population increased, the
racial profile changed, but the distribution of income was reasonably stable,
with a higher amount of the population in the lowest income range and very few
people in the highest income range. Indeed, the income  difference is
significant when compared to the city-wide distribution.

\begin{figure}
    \centering 
    \includegraphics[width=0.95\linewidth]{laBH.pdf}
    \caption{Top: Geographic changes in the majority Black population cluster
    (orange) and Laborers cluster (green). Bottom: Income evolution for this
    region (black) and the whole city (gray).\label{fig:labh}}
\end{figure}

A portion of this region is classified into the green cluster in 2010,
indicating a majority white population. To further understand that change, we
clear the current selection, and select all regions that changed from orange in
1970 to green in 2010, using the transition matrix. A portion of the resulting
region, near the Florence-Graham region, is depicted in Figure~\ref{fig:laVol},
along with the temporal evolution of Race. Despite this difference, the other
aspects are similar to the ones from the region in Figure~\ref{fig:labh}, with
slightly lower income and education profiles. While the racial aspect changed
considerably, the economic and educational aspects stayed the same.

\begin{figure}
    \centering 
    \includegraphics[width=0.95\linewidth]{laVol.pdf}
    \caption{Details for a volatile region contained in the area of
         Figure~\ref{fig:labh}. This region went from Black to Hispanic to
        White.\label{fig:laVol}}
\end{figure}

While Toronto is more dynamic than Los Angeles, possibly due to size
differences, the volatile regions shown in Figure~\ref{fig:toVol} did not change
as quickly or dramatically as the ones shown in Figure~\ref{fig:laVol}, which
involved twice as many people. We found this trend to be related to the
countries themselves, Canadian cities have larger areas undergoing slow, gradual
changes, whereas American cities have more general stability, but quicker
changes in smaller scales. The supplementary material contains brief summaries
of all the regions accessible in this prototype.

\section{Expert feedback}
%This could be a big step toward better understanding of evolving urban morphology.
To assess the proposed method, we contacted academic and industry experts in
sociology and urban sciences. The supplementary material contains all our
communication with five experts, identified from A to E.

Their response was positive,  mentioning that the prototype allows them to
analyze census data without the additional work of obtaining and cleaning the
data (A, B, E), and it allows the inclusion of geographic visual analysis tools
in their research process (D). It enables the users to tell different stories
about neighborhoods/cities and their changes (A), visualize the relationship
between key urban variables over time (D), offering a quick way to identify
particular neighborhoods that one may be interested in studying more in depth
around a particular issue or efficiently understanding the context of an area
(E).  Indeed, the experts identified gentrification processes in Manhattan (B)
and Dallas (E), reinforced a hypothesis for occupational clustering (D), and
highlighted how the method can be used to compare neighborhoods and cities (A).
These remarks attest that we satisfied the design requirements regarding the
existing demographic groups and their changes.

While the interface was "easy to navigate" (B), it was also considered
"overwhelming" (A), "intimidating" (E), and "tricky to interpret" (C), possible
side-effects of our effort to increase  representational accuracy, where we
avoided using simplified labels. Specifically, while identifying clusters by
their most relevant variables was welcome, the overlap of information from
different clusters in the boxplot was "a bit confusing" (C). While the map of
trajectories was mentioned as a "good summary map", how it related to the
clustering method was unclear (C). While we provide different options on how the
colors are used on this map, both are sub optimal, reliably representing several
distinct entities using colors is still an open problem in visualization.

The experts also mentioned the poor responsiveness of the method when changes in
the clustering parameters required server-side processing (B,D). Indeed, the
current implementation can take a few minutes to cluster regions with high
number of CTs, like Los Angeles or Brooklyn.

Most of the experts demonstrated interest in using our method in their research
(A, B, D, E), aiming to use the census data as a backdrop for other datasets,
providing demographic context. They also mentioned the need to export subsets of
data, plots, and maps to be used in reports and publications (C, D, E).

More importantly, while these experts were aware that our method does not
perform geographic normalization, none of them mention it. We did not
specifically ask if this difference led to unexpected results, but rather
inquired if they found interesting insights, leading them to consider the
interface, its options, and results as a whole. Therefore the method was
neutrally compared against other, traditionally normalized results. We interpret
the fact that most of them were interested in the next steps as an affirmation
of the accuracy of the method.