### 2019-06-13 paper/related.tex

\section{Related Work}
\label{sec:related}

Since our problem encompasses several fields, we divide this section into
specific sub problems: \emph{longitudinal demographic studies}, describing the
traditional tabular approach to longitudinal studies; \emph{data
representation}, elaborating how evolving geographic data can be represented for
processing; \emph{data clustering}, briefly reviewing existing clustering
methods; and \emph{cluster characterisation}, articulating how clusters can be
visually summarised.

\subsection{Longitudinal demographic studies}
Census data is used not only to discover demographic
patterns~\citep{Firebaugh2016}, but to correlate demographic characteristics to
other measurements~\citep{diez1997neighborhood}. However, longitudinal studies
are rare, because they are difficult : \emph{"(...) One of the most challenging
and fascinating areas in spatial statistics is the synthesis of spatial data
collected at different spatial scales(...)"}~\citep{gotway2002combining}. While
census tract level data is readily available for the US since at least
1910~\citep{nhgis}, most studies consider the period between 1970 and 2010,
using pre-harmonised data from the Longitudinal Tract Data
Base~\citep{Logan2014}. Despite its inherent
errors~\citep{Logan2016,Hallisey2017}, this dataset has become the standard
source for longitudinal demographic data at the neighbourhood scale, with
similar efforts appearing in other countries~\citep{Liu2015,Lee2015,Allen2018}.
These datasets have been highly significant for the field. Yet they also limit
the universe of  data that can be used to study neighbourhood change, since any
new datasets would need to be similarly processed in order to be rendered
compatible with these sources.

Another option considers the use of grid
data~\citep{Dmowska2017,Dmowska2018,stepinski2019imperfect}. Beyond the
increased spatial accuracy, this approach does not require complex harmonisation
when new data is considered, if the grids are compatible. However, demographic
data is usually not available in this format, especially from older sources.
Additionally, the conversion from tabulation areas can introduce significant
errors.

Given these challenges, it is worth considering new alternatives. In this work,
we propose a novel methodology that entirely avoids the problems of geographical
harmonisation, considering each measurement using its actual geographic region.
It does not require regions to be consistent across time because they are
naturally represented as different entities.

\subsection{Data representation}

Network based representation of geographic information is fairly well explored
in the literature, as a basis for topological methods for event
detection~\citep{Doraiswamy2014}, leveraging signal processing on
graphs~\citep{shuman2013emerging,sandryhaila2013discrete} to find patterns and
outliers~\citep{Valdivia2015,Dias2015,Alce2018}. Networks are well suited to
represent trajectories as
well~\citep{VonLandesberger2016,Huang2016,chen2015survey}, allowing the use of
graph visualisation methods~\citep{Vehlow2015,Beck2014}. Our proposed method
builds upon this literature. We leverage a network-based representation that
removes the requirement for consistency in the measurement regions. Each region
in time corresponds to a different node. Instead of a collection of time-series,
the data is represented as a dynamic network.

Networks have been used to represent census data for clustering purposes
~\citep{Dias2015,Setiadi2017}, but these works did not explore temporal
evolution, where they are particularly powerful. Networks allow a natural
representation of these inconsistent regions, with both spatial and temporal
connections. There are other possible representations that have similar
properties, but we adopted networks to allow the use of the vast existing
literature and methods.

\subsection{Data clustering}

Data clustering is one of the elementary processes for data analysis,
simplifying the data into a smaller number of homogeneous sets that can be
interpreted in the same way. There is no shortage of contributions for this
problem~\citep{Fahad2014}, but most neighbourhood related applications still
rely on k-means~\citep{jain2010data,Delmelle2016} and, to a lesser extent, Self
Organising Maps~\citep{Delmelle2017,Ling2016}.

However, a method for geographic data analysis should not ignore the geographic
component of the data, and we extend research exploring ways to incorporate it.
One straightforward option for agglomerative methods~\citep{han2001spatial} is
to consider only nearby clusters for merging~\citep{Chavent2017}, which can also
be done for k-means~\citep{soor2018extending}. Alternatively, the spatial
distance could be directly added to the inter-cluster metric~\citep{Chavent2017}
via a mixing parameter, which adds flexibility to the method, but introduces the
problem of finding the correct application-dependent values.

Indeed, one crucial step in most clustering algorithms is the definition of the
number of clusters. We avoid this problem by considering hierarchical
methods~\citep{soille2012morphological}, where the result is not a partition of
the data, but a tree of partitions, similar to a dendogram. This approach is
interesting for interactive methods, because it allows the user to change the
number of clusters on the fly. Since our data is represented as a network, we
opted for an heuristic variation of the maximum weighted matching algorithm
called \emph{sorted maximal matching}~\citep{markus2017}, which merges clusters
based on the weights of the edges between pairs of clusters.

\subsection{Cluster characterisation}
While visualisation has gained prominence as a crucial component of scientific
discovery, justification, and communication \todo{Add Cites: maybe Gelman?
Tufte?}, visually representing evolving spatial data is a challenging old
problem~\citep{monmonier1990strategies,andrienko2003exploratory,ferreira2015visual,Zheng2016}.

Most geographic data is naturally bidimensional and maps work well in this
case~\citep{Zheng2016,ward2015interactive}, but the additional temporal
dimension cannot be so naturally represented. One straightforward option is to
leverage tridimensional plots~\citep{andrienko2014visualization,Tominski2012a},
but this can lead to visual obstructions or scaling problems unless a
tridimensional display device is used. A simpler, well adopted,
option is to display a map that corresponds to a subset of the temporal
information, allowing the user to change the time with an associated
control~\citep{Chen2017,Valdivia2015,Alce2018,Doraiswamy2014}. Small multiples
can be used~\citep{VonLandesberger2016}, but only when there are few temporal
snapshots. However, none of these options is suitable to represent many
variables at the same time.

Using data clustering, we can represent the region's cluster instead of all the
its variables~\citep{Alce2018,Valdivia2015,VonLandesberger2016}. While this
simplifies the geographic portion of the visualization, it introduces the
problem of how to summarise the contents of each cluster. One traditional
approach is to use parallel coordinates plot~\citep{ferreira2015urbane}, but
these they can get cluttered representing similar clusters over several
variables. Further, for demographic applications, the clusters are usually
strongly characterised by a small subset of
values~\citep{Delmelle2016,Delmelle2017}. Therefore, in the proposed method, we
identify the variables that are most relevant to the characterisation of each
cluster. The distribution of values on that variable is then represented using a
boxplot, a well known statistical plot displaying basic properties of the
distributions.